---
title: "Linear Fit For X Squared"
author: "Clark Gaylord"
date: "11/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Linear regression can be used regardless of whether the data are in fact linear. Sometimes the assumption of linearity results in particularly poor results, especially if there is a need for extrapolation. However, the linear model is also a "first order approximation" and sometimes the simplicity of the model itself brings its own virtues.

Let's consider how well (or how poorly) a linear model might work for a scenario in which we *know* that our data are not linear. Consider the situation where we have y = x^2, say for x between 1 and 10. Although clearly x^2 is not linear, it is still the case that low values of x imply low values of y, and high values of x imply high values of y. So let's see how well this works.

```{r}
xy <- data.frame(x = seq(1,10))
xy$y <- xy$x^2
plot(y~x, data=xy, main="Y = X^2")
```

We wish to introduce the "linear model" command in R: lm().

```{r}
xy.lm <- lm(y~x, data=xy)
xy.lm
summary(xy.lm)
```

Note that the R-squared value is 95%, which is generally considered to be quite high.

```{r}
plot(y~x, data=xy,
     ylim = c(-25,105),
     main = "Fitting a line to Y=X^2")
abline(xy.lm)
text(x=8,y=50,pos=4,label="y = 11 x - 22")
```

We see that while there is clearly a systematic misrepresentation of the model to the data, the line does somewhat approximate the data.